---
title: "CCLE new"
output: html_notebook
---

# packages
```{r warning = FALSE, message = FALSE}
setwd("C:/Users/Sohyeon Kim/add_decomp_linux")

library(dplyr)
library(stringr)
library(Matrix)
library(data.table)
library(Rcpp)
library(splines)
library(fda)
library(glmnet)
library(foreach)
library(doParallel)
library(gplots)
library(tibble)

source("https://raw.githubusercontent.com/S0Hye0NKim/add_decomp/master/functions/add_decomp_function.R")
sourceCpp("[KSH]add_decomp_function.cpp")
```

# 0. Data & Screening


## 0-(1) Screening Y


```{r message = FALSE, warning = FALSE}
Y <- fread("CCLE_Y_new.csv")

m <- ncol(Y)

Y_sd <- Y[, lapply(.SD, sd)] %>% as.matrix %>% t()
data.frame(Y_sd) %>%
  mutate(group = rownames(.)) %>%
  dplyr::select(group, Y_sd) %>%
  arrange(Y_sd) %>% head(5)

Y <- Y[, -"Panobinostat"] %>%
  column_to_rownames(var = "CCLE_Name") %>%
  as.matrix
m <- ncol(Y)
```



Panobinostat 의 sd = 0.08 -> 제거

## 0-(2) Screening X

### 0-(2)-(1) Standard deviation of X

 * Assumption : Genes with small sd are irrelevant to the drug responses. 

```{r}
X <- fread("CCLE_X_new.csv")
X <- X[, -"Broad_ID"]

X_sd <- X[, -"CCLE_Name"][, lapply(.SD, FUN = sd)] %>%
  as.matrix() %>% t()
idx_scr_sd <-  data.frame(X_sd) %>%
  mutate(variable = rownames(.)) %>%
  dplyr::select(variable, X_sd) %>%
  arrange(-X_sd) %>% 
  .[1:10000, ] %>% .$variable

X <- dplyr::select(X, c(CCLE_Name,all_of(idx_scr_sd))) %>%
  column_to_rownames(var = "CCLE_Name")
```



### 0-(2)-(2) Correlation of X and Y

$$\frac{1}{23}\sum_{g\ne11}|\text{cor}(x_j, y_g)|$$

 * Assumption : Genes having low correlations with drugs may not affect the drug responses. 

```{r}
n <- nrow(X)
p <- 100


idx_scr_cor <- cor(X, Y) %>% abs %>%
  apply(MARGIN = 1, FUN = mean) %>%
  data.frame(variable = names(.), cor = .) %>%
  arrange(-cor) %>% first(p) %>% .$variable

X <- dplyr::select(X, all_of(idx_scr_cor)) %>% 
  apply(MARGIN = 2, FUN = function(x) (x - mean(x))/(sd(x)*sqrt((n-1)/n))) %>%
  cbind(1, .)
colnames(X)[1] <- "intercept"
gene_ex <- colnames(X)
```


## 0-(3). spline

```{r}
K <- 5
b <- 15
tau_seq <- seq(from = 0.35, to = 0.65, length.out = b)
idx_tau <- (tau_seq >= "0.4" & tau_seq <= "0.6")
tau_seq_real <- tau_seq[idx_tau]

knots_seq <- seq(min(tau_seq) - 0.02, max(tau_seq) + 0.02, length.out = K)
Phi <- fda::bsplineS(tau_seq, breaks= knots_seq, norder=2, nderiv=0, returnMatrix=FALSE)

V <- calc_V(as.matrix(X), Phi)
```


## 0-(4). first initial value 

```{r}
set.seed(1)
X <- as.matrix(X)
Y <- as.matrix(Y)

 lasso_coef <- matrix(nrow = p+1, ncol = m)
  for(g in 1:m) {
    cv.lasso <- cv.glmnet(x = X[, -1], y = Y[, g], alpha = 1, type.measure = "mae")
    lasso_model <- glmnet(X[, -1], Y[, g], family = "gaussian", alpha = 1, lambda = cv.lasso$lambda.min)
    lasso_coef[, g] <- c(lasso_model$a0, as.vector(lasso_model$beta))
  }

theta_init <- matrix(nrow = (p+1)*K, ncol = m)
  for(g in 1:m) {
    for(j in 0:p) {
      theta_init[((j*K)+1):((j+1)*K), g] <- lasso_coef[j+1, g]
    }
  }
  
  Y_modified <- Y - X%*%lasso_coef
  ridge_coef <- matrix(nrow = p+1, ncol = m)
  for(g in 1:m) {
    cv.ridge <- cv.glmnet(x = X[, -1], y = Y_modified[, g], alpha = 0, type.measure = "mae")
    ridge_model <- glmnet(X[, -1], Y_modified[, g], family = "gaussian", alpha = 0, lambda = cv.ridge$lambda.min)
    ridge_coef[, g] <- c(ridge_model$a0, as.vector(ridge_model$beta))
  }
  alpha_init <- ridge_coef
```



# 1. Initial value

```{r}
init_val <- add_decomp(delta = 1, lambda_1 = 0.01, lambda_2 = 0.001, tol_error = 0.1^5, max_iter = 50,
                         X = X, Y = Y, V = V, Phi = Phi, 
                         theta_0 = theta_init, Z_0 = X%*%alpha_init, tau_seq = tau_seq, weight = FALSE)

```

```{r}
init_val$Z %>% rankMatrix
gamma_tau_hat <- est_gamma(Phi[idx_tau, ], init_val$theta)

check_sp_table(true = matrix(0, nrow = (p+1), ncol = m), 
               est = gamma_tau_hat, tau_seq = tau_seq_real, table = TRUE)
```



# 2. BIC

```{r}
log_lamb1 <- c( seq(0.1, 2, length.out = 20))
lamb1_seq <- exp(log_lamb1)

log_lamb2 <- c(seq(1, 2, length.out = 20))
lamb2_seq <- exp(log_lamb2)

BIC_table <- list()

for(idx in 1:length(lamb1_seq)) {
  lamb1 <- lamb1_seq[idx]
  cl <- makeCluster(20) #not to overload your computer
  registerDoParallel(cl) # Ready to parallel
  
  temp_BIC <- foreach(lamb2 = lamb2_seq, .noexport = "add_decomp") %dopar% {
    library(dplyr)
    library(splines)
    library(Matrix)
    library(glmnet)
    library(fda)
    
    BIC_simul <- add_decomp_BIC(X, Y, V, Phi, theta_0 = init_val$theta, Z_0 = init_val$Z, tau_seq, tau_seq_real, delta = 1, 
                        lamb1_seq = lamb1, lamb2_seq = lamb2, max_iter = 50)
    BIC_simul$table
  }
  stopCluster(cl)
  BIC_table[[idx]] <- temp_BIC
}
 

BIC_table %>% lapply(FUN = function(x) bind_rows(x)) %>%
  bind_rows() %>%
  mutate(LR_part = r_hat * max(1, m) / (2*n*m),
         S_hat_net = S_hat - num_nz_intercept,
         LR = log(p) * log(log(n)) * LR_part, 
         SP = log(p) * log(log(n)) * K * S_hat_net / (2*n*m),
         BIC = log_Q + LR + SP) %>%
  mutate_all(as.numeric) %>%
  filter(S_hat_net != 0) %>%
  arrange(BIC) %>%  
  head(6)

```


# 3. Proposed Model

```{r}
log_lamb1 <- c( seq(0.1, 2, length.out = 20))
lamb1_seq <- exp(log_lamb1)

log_lamb2 <- c(seq(1, 2, length.out = 20))
lamb2_seq <- exp(log_lamb2)

lambda_1 <- lamb1_seq[10]
lambda_2 <- lamb2_seq[20]
```



```{r}
result <- add_decomp(delta = 1, lambda_1 = lambda_1, lambda_2 = lambda_2, tol_error = 0.1^5, max_iter = 50,
                         X = X, Y = Y, V = V, Phi = Phi, 
                         theta_0 = init_val$theta, Z_0 = init_val$Z, tau_seq = tau_seq, weight = TRUE)
```



## Low-rank matrix

```{r}
X_tilde <- X[, -1]

L_hat <- solve(t(X_tilde) %*% X_tilde) %*% t(X_tilde) %*% result$Z
L_hat %>% rankMatrix
```


```{r}
sing_val <- svd(L_hat)$d
sum(sing_val[1:1])/sum(sing_val)
rankMatrix(result$Z)
sing_val %>% round(2)
```



```{r}
colnames(L_hat) <- colnames(Y)
heatmap.2(L_hat, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none")
title("L hat heatmap")
```

```{r}
r_hat <- rankMatrix(L_hat)[1]
L_hat_t <- t(L_hat)
dist_mat <- dist(L_hat_t, method = "euclidean")
cluster <- hclust(dist_mat, method = "complete")
plot(cluster, cex = 0.6, main = "clustering from L hat")
rect.hclust(cluster, k = r_hat, border = 2:(1+r_hat))
```


```{r}
SVD_L <- svd(L_hat)
col_sp <- SVD_L$v %>% .[, 1:3]

data.frame(col_sp) %>%
  `colnames<-`(value = paste0("V", 1:3)) %>%
  mutate(treatment = colnames(Y)) %>%
  gather(key = "Variable", value = "value", -treatment) %>%
  filter(Variable != "V1") %>%
  ggplot() +
  geom_col(aes(x = treatment, y = value)) +
  theme(axis.title.x = element_blank(), 
    axis.text.x = element_text(angle = 90)) +
  facet_wrap(~Variable)
```



```{r}
weak_signals <- c("Lapatinib", "Nutlin-3", "Paclitaxel", "PHA-665752", "Irinotecan", "17-AAG", "LBW242", "L-685458", "PD-0332991", "TAE684", "Topotecan", "AZD0530", "Erlotinib", "Nilotinib", "PF2341066", "Sorafenib")

L_hat[, weak_signals] %>%
  apply(MARGIN = 2, FUN = function(x) abs(x) %>% median)

```


```{r}
colnames(result$Z) <- colnames(Y)
rownames(result$Z) <- rownames(Y)
heatmap.2(result$Z, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none")
title("Z hat heatmap")
```

```{r}
r_hat <- rankMatrix(result$Z)[1]
Z_hat_t <- t(result$Z)
dist_mat <- dist(Z_hat_t, method = "euclidean")
cluster <- hclust(dist_mat, method = "complete")
plot(cluster, cex = 0.6, main = "clustering from Z hat")
rect.hclust(cluster, k = r_hat, border = 2:(1+r_hat))
```


## Sparse matrix

```{r}

gamma_tau_hat <- est_gamma(Phi[idx_tau, ], result$theta)
```



```{r}
zero_idx_est <- lapply(gamma_tau_hat, FUN = function(x) which(abs(x) < 0.1^5, arr.ind = TRUE) %>% as_tibble) %>%
    bind_rows (.id = "tau") %>%
    group_by(row, col) %>%
    summarise(zero = n(), .groups = "keep") %>%
    filter(zero == length(gamma_tau_hat))

Neg <- nrow(zero_idx_est)
Pos <- prod(dim(gamma_tau_hat[[1]])) - Neg

data.frame("Positive"=Pos, "Negative" = Neg)


```

```{r}
library(ggplot2)
Y_col <- factor(colnames(Y), ordered = TRUE, levels = colnames(Y))

lapply(gamma_tau_hat, FUN = function(x) data.frame(col = rep(1:m, each = (p+1)),
                                                   row = rep(1:(p+1), m),
                                                   value = as.vector(x))) %>%
  `names<-`(tau_seq_real) %>%
  bind_rows(.id = "tau") %>%
  mutate(is_nonzero = ifelse(abs(value) < 0.1^5, 0, 1)) %>%
  group_by(col, row) %>%
  summarise(is_nonzero = mean(is_nonzero)) %>%
  ungroup() %>%
  mutate(is_nonzero = ifelse(is_nonzero == 0, 0, 1) %>% as.character, 
         treatment = rep(Y_col, each = p+1)) %>%
  ggplot(aes(x = treatment, y = -row, fill = is_nonzero)) +
  geom_tile(color = "grey") +
  scale_fill_manual(values = c("white", "black")) + 
  theme_bw() +
  theme(axis.line = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    axis.title.x = element_blank(), 
    axis.text.x = element_text(angle = 90), 
    axis.title.y = element_blank(), 
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank()) +
  ggtitle(expression("S("~tau[l]~") in AD model"))
```


## smooth function


```{r}
library(tidyr)
library(rebus)

idx_nz_est <- lapply(gamma_tau_hat, FUN = function(x) which(abs(x) > 0.1^5, arr.ind = TRUE) %>% as_tibble) %>%
    bind_rows (.id = "tau") %>%
    group_by(row, col) %>%
    summarise(zero = n()) 

gamma_data <- gamma_tau_hat %>%
  lapply(FUN = function(x) x[idx_nz_est$row, idx_nz_est$col] %>% diag %>%
           data.frame(row = idx_nz_est$row, col = idx_nz_est$col, value = . )) %>%
  `names<-`(value = tau_seq_real) %>%
  bind_rows(.id = "tau") %>%
  filter(row != 1, col == 13) %>% 
  unite(col = index, row, col, sep = ",") %>%
  mutate(index = paste0("(", index, ")"), tau = as.numeric(tau))

row_pattern <- "\\(" %R% one_or_more(DGT)
col_pattern <- one_or_more(DGT) %R% "\\)"

gamma_data <- mutate(gamma_data, row = str_extract(index, pattern = row_pattern) %>% parse_number, 
                     col = str_extract(index, pattern = col_pattern) %>% parse_number) %>%
  left_join(data.frame(row = 1:ncol(X), row_names = colnames(X)), by = "row") %>%
  left_join(data.frame(col = 1:ncol(Y), col_names = colnames(Y)), by = "col")

gamma_data_list <- split(gamma_data, gamma_data$index)
spline_int <- lapply(gamma_data_list, FUN = function(x) as.data.frame(spline(x$tau, x$value))) %>%
  bind_rows(.id = "index") %>%
  mutate(row = str_extract(index, pattern = row_pattern) %>% parse_number, 
         col = str_extract(index, pattern = col_pattern) %>% parse_number) %>%
  left_join(data.frame(row = 1:ncol(X), row_names = colnames(X)), by = "row") %>%
  left_join(data.frame(col = 1:ncol(Y), col_names = colnames(Y)), by = "col")

ggplot(gamma_data) +
  geom_point(aes(x = tau, y = value, color = row_names)) +
  geom_line(data = spline_int, aes(x = x, y = y, group = row_names, color = row_names)) +
  ggtitle(expression("S("~tau[l]~") smooth function"))
```

# 4. LR model

```{r}
set.seed(1)
ridge_coef <- matrix(nrow = p+1, ncol = m)
for(g in 1:m) {
  cv.ridge <- cv.glmnet(x = X[, -1], y = Y[, g], alpha = 0, type.measure = "mae")
  ridge_model <- glmnet(X[, -1], Y[, g], family = "gaussian", alpha = 0, lambda = cv.ridge$lambda.min)
  ridge_coef[, g] <- c(ridge_model$a0, as.vector(ridge_model$beta))
}

first_init_LR <- ridge_coef
init_val_LR <- LR_model_r(delta = 1, lambda = 0.1, tol_error = 0.1^5, max_iter = 50, 
                          X = X, Y = Y, Z_0 = X %*% first_init_LR, tau_seq = tau_seq, weight = FALSE)
```


```{r}
lamb_seq <- seq(0.1^20, 0.1, length.out = 25)
r_X <- rankMatrix(X[, -1])
BIC_simul <- LR_model_BIC(X = X, Y = Y, Z_0 = init_val_LR$Z, tau_seq = tau_seq, tau_seq_real = tau_seq_real, delta = 1, lamb_seq = lamb_seq, max_iter = 50, r_X = 1)
  
BIC_params <- BIC_simul$min_BIC %>%
  arrange(BIC_log_p) %>%
  mutate_all(as.numeric) %>%
  head(1)
```


```{r}
lamb_seq <- seq(0.1^20, 0.1, length.out = 25)
lambda <- lamb_seq[16]

result_LR <- LR_model_r(delta = 1, lambda = lambda, tol_error = 0.1^5, max_iter = 50, X = X, Y = Y, Z_0 = init_val_LR$Z, tau_seq = tau_seq, weight = TRUE)
```



## Low-rank matrix

```{r}
X_tilde <- X[, -1]

L_hat <- solve(t(X_tilde) %*% X_tilde) %*% t(X_tilde) %*% result_LR$Z
L_hat %>% rankMatrix
```


```{r}
sing_val <- svd(L_hat)$d
sum(sing_val[1:1])/sum(sing_val)
rankMatrix(result_LR$Z)
sing_val %>% round(2)
```



```{r}
colnames(L_hat) <- colnames(Y)
heatmap.2(L_hat, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none")
title("L hat heatmap")
```

```{r}
r_hat <- rankMatrix(L_hat)[1]
L_hat_t <- t(L_hat)
dist_mat <- dist(L_hat_t, method = "euclidean")
cluster <- hclust(dist_mat, method = "complete")
plot(cluster, cex = 0.6, main = "clustering from L hat")
rect.hclust(cluster, k = r_hat, border = 2:(1+r_hat))
```


```{r}
colnames(result_LR$Z) <- colnames(Y)
rownames(result_LR$Z) <- rownames(Y)
heatmap.2(result_LR$Z, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none")
title("Z hat heatmap")
```

```{r}
r_hat <- rankMatrix(result_LR$Z)[1]
Z_hat_t <- t(result_LR$Z)
dist_mat <- dist(Z_hat_t, method = "euclidean")
cluster <- hclust(dist_mat, method = "complete")
plot(cluster, cex = 0.6, main = "clustering from Z hat")
rect.hclust(cluster, k = r_hat, border = 2:(1+r_hat))
```


# 5. SP model

```{r}
set.seed(1)
lasso_coef <- matrix(nrow = p+1, ncol = m)
for(g in 1:m) {
  cv.lasso <- cv.glmnet(x = X[, -1], y = Y[, g], alpha = 1, type.measure = "mae")
  lasso_model <- glmnet(X[, -1], Y[, g], family = "gaussian", alpha = 1, lambda = cv.lasso$lambda.min)
  lasso_coef[, g] <- c(lasso_model$a0, as.vector(lasso_model$beta))
}
  
first_init_SP <- matrix(nrow = (p+1)*K, ncol = m)
for(g in 1:m) {
  for(j in 0:p) {
    first_init_SP[((j*K)+1):((j+1)*K), g] <- lasso_coef[j+1, g]
  }
}
  
init_val_SP <- SP_model_r(delta = 1, lambda = 0.01, tol_error = 0.1^5, max_iter = 50, 
                          X = X, Y = Y, V = V, Phi = Phi, theta_0 = first_init_SP, tau_seq = tau_seq, weight = FALSE)
```

```{r}
idx_tau <- (tau_seq >= 0.4 & tau_seq <= 0.6)
gamma_tau_hat_SP <- est_gamma(Phi[idx_tau, ], init_val_SP$theta)

check_sp_table(true = matrix(0, nrow = (p+1), ncol = m), est = gamma_tau_hat_SP, table = TRUE, tol = 0.1^5, tau_seq = tau_seq_real)
```


```{r}
log_lamb <- c(seq(1, 2, length.out = 20))
lamb_seq <- exp(log_lamb)

BIC_table <- list()
cl <- makeCluster(20) #not to overload your computer
registerDoParallel(cl) # Ready to parallel
  
BIC_table <- foreach(lambda = lamb_seq, .noexport = "add_decomp") %dopar% {
  library(dplyr)
  library(splines)
  library(Matrix)
  library(glmnet)
  library(fda)
    
  BIC_simul <- SP_model_BIC(X, Y, V, Phi, theta_0 = init_val_SP$theta,
                            tau_seq, tau_seq_real, lamb_seq = lambda, max_iter = 50, delta = 1)
    BIC_simul$BIC_data
}
stopCluster(cl)
  
BIC_params <- BIC_table %>%
  bind_rows() %>%
  filter(S_hat_net != 0) %>%
  arrange(BIC_log_p) %>%
  mutate_all(as.numeric) %>%
  head(1)
```

```{r}
log_lamb <- c(seq(1, 2, length.out = 20))
lamb_seq <- exp(log_lamb)

lambda <- lamb_seq[19]
result_SP <- SP_model_r(delta = 1, lambda = lambda, tol_error = 0.1^5, max_iter = 50, X = X, Y = Y, V = V, Phi = Phi, theta_0 = init_val_SP$theta, tau_seq = tau_seq, weight = TRUE)
```

```{r}
idx_tau <- (tau_seq >= 0.4 & tau_seq <= 0.6)
gamma_tau_hat_SP <- est_gamma(Phi[idx_tau, ], result_SP$theta)

check_sp_table(true = matrix(0, nrow = (p+1), ncol = m), est = gamma_tau_hat_SP, table = TRUE, tol = 0.1^5, tau_seq = tau_seq_real)
```

```{r}
library(ggplot2)
Y_col <- factor(colnames(Y), ordered = TRUE, levels = colnames(Y))

lapply(gamma_tau_hat_SP, FUN = function(x) data.frame(col = rep(1:m, each = (p+1)),
                                                   row = rep(1:(p+1), m),
                                                   value = as.vector(x))) %>%
  `names<-`(tau_seq_real) %>%
  bind_rows(.id = "tau") %>%
  mutate(is_nonzero = ifelse(abs(value) < 0.1^5, 0, 1)) %>%
  group_by(col, row) %>%
  summarise(is_nonzero = mean(is_nonzero)) %>%
  ungroup() %>%
  mutate(is_nonzero = ifelse(is_nonzero == 0, 0, 1) %>% as.character, 
         treatment = rep(Y_col, each = p+1)) %>%
  ggplot(aes(x = treatment, y = -row, fill = is_nonzero)) +
  geom_tile(color = "grey") +
  scale_fill_manual(values = c("white", "black")) + 
  theme_bw() +
  theme(axis.line = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    axis.title.x = element_blank(), 
    axis.text.x = element_text(angle = 90), 
    axis.title.y = element_blank(), 
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank()) +
  ggtitle(expression("S("~tau[l]~") in SP model"))
```

```{r}
lapply(gamma_tau_hat_SP, FUN = function(x) data.frame(col = rep(1:m, each = (p+1)),
                                                   row = rep(1:(p+1), m),
                                                   value = as.vector(x))) %>%
  `names<-`(tau_seq_real) %>%
  bind_rows(.id = "tau") %>%
  mutate(is_nonzero = ifelse(abs(value) < 0.1^5, 0, 1)) %>%
  group_by(col, row) %>%
  summarise(is_nonzero = mean(is_nonzero), avg = mean(value)) %>%
  ungroup() %>%
  mutate(is_nonzero = ifelse(is_nonzero == 0, 0, 1) %>% as.character, 
         treatment = rep(Y_col, each = p+1)) %>%
  left_join(y = data.frame(row = 1:101, gene_ex = colnames(X)), by = "row") %>%
  filter(row != 1, is_nonzero == 1) %>%
  dplyr::select(-is_nonzero)
```



