---
title: "Toy Example"
author: "Sohyeon Kim"
date: "7/21/2020"
output: 
  github_document:
    pandoc_args: --webtex
header-includes:
  - \usepackage{kotex}
---

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(splines)
library(Matrix)
```


# 0. Initial Setting

$$\begin{aligned}Y &\in \mathbb{R}^{n\times g}\\Y &= XB(\tau) + \epsilon(\tau) , \; X\sim N(0, 1)\\ X &\in\mathbb{R}^{n \times (p+1)}, \; \;B(\tau) \in \mathbb{R}^{(p+1)\times g} , \; and \; Q_\tau(\epsilon(\tau)|X) = 0\end{aligned}$$

$$\begin{aligned}\beta_0(\tau) &= \beta_0 + Q_\tau(\epsilon) \\ \epsilon(\tau) &= \epsilon- Q_\tau(\epsilon), \; \epsilon \sim N(0, 0.1^2)\end{aligned}$$

 * n : # of observation. 
 * m : # of group. 
 * p : # of covariate. 
 * r : nonzero entry in true $B(\tau)$
 * K : # of basis function.
 * b : # of regional quantiel we consider.
 
For toy example, I will use n = 300, p = 50, g = 20.

```{r}
set.seed(0)
n <- 300
m <- 20
p <- 50
r <- 500
b <- 20

X <- matrix(rnorm(n*p, mean = 0, sd = 1), nrow = n) %>% cbind(1, .)   #add intercept term in X
sp_mat <- rsparsematrix(nrow = p+1, ncol = m, nnz = r)             # make sparse matrix
LR_mat <- matrix(rnorm(m*(p+1), mean = 0, sd = 1), ncol = m)      # make low rank matrix using SVD
SVD <- svd(LR_mat)
D_mat <- diag(SVD$d)
idx <- sample(1:m, m/2)
D_mat[idx, idx] <- 0
LR_mat <- SVD$u %*% D_mat %*% t(SVD$v)          

true_B <- sp_mat + LR_mat            # B(tau) = sparse matrix + low rank matrix

eps <- matrix(rnorm(n*m, mean = 0, sd = 0.1), nrow = n)
#tau <- 0.5
#true_B_tau <- true_B
#true_B_tau[1, ] <- true_B[1, ] + qnorm(tau, mean = 0, sd = 0.1)
#eps_tau <- eps - qnorm(tau, mean = 0, sd = 0.1)

Y <- X%*%true_B + eps
```

$B(\tau)$ = sparse matrix + low rank matrix


#### Question

 1. Y를 generate할 때, 사용하는 $B(\tau)$ 값에 tau depend? 
 2. 우선 $\tau = 0.5$ 로 generate 시켰는데, 이렇게 하는게 맞나? 

# 1. Preliminary
 
### 1-1. Check function

```{r}
check_ft <- function(x, tau) {
  z <- ifelse(x<0, (tau-1)*x, tau*x)
  return(z)
}
```

$$\rho_\tau(x) = (\tau - \mathbb{I}(x<0))x$$

### 1-2. Basis function

```{r}
K <- 15

tau_seq <- seq(from = 0.25, to = 0.75, length.out = b) 
Phi <- bs(tau_seq, df = K, degree = 3, intercept = TRUE)
```


$$\begin{aligned}\boldsymbol{v}_i^{(\ell)} &= (\boldsymbol{x}_i \otimes \Phi(\tau_{\ell}))^T\\ &=(\Phi(\tau_\ell)^T, x_{i1}\Phi(\tau_\ell)^T, \dots, x_{ip}\Phi(\tau_\ell)^T) \\ &\in \mathbb{R}^{(p+1)K}\end{aligned}$$

### 1-3. New design matrix for regional quantile

```{r}
V <- list()
for(l in 1:b){
  V_l <- matrix(nrow = n, ncol = (p+1)*K)
  for(i in 1:n) {
    v_i_l <- kronecker(X[i, ], Phi[l, ])
    V_l[i, ] <- v_i_l 
  }
  V[[l]] <- V_l
}
```


 * Dimension for V : $b \times n \times (p+1)K$

$$\boldsymbol{V}^{(\ell)} = \begin{bmatrix} v_1^{(\ell)} \\v_2^{(\ell)} \\ \vdots \\ v_n^{(\ell)}\end{bmatrix} \in \mathbb{R}^{n\times (p+1)K}$$

### 1-4. Initial value for estimator


$$\boldsymbol{\eta}^{(g)}, \boldsymbol{\theta}^{(g)}, \boldsymbol{w}^{(g)}\in \mathbb{R}^{(p+1)\times K}\\ \boldsymbol{\alpha}^{(g)} \in \mathbb{R}^{p+1} \\ \boldsymbol{e}^{(\ell)(g)}, \boldsymbol{u}^{(\ell)(g)}, \in \mathbb{R}^{n}\\subject\; to \; \boldsymbol{\eta}^{(g)} - \boldsymbol{\theta}^{(g)} = 0, \; and \;\boldsymbol{Y}^{(g)} - \boldsymbol{X}\boldsymbol{\alpha}^{(g)} - \boldsymbol{V}^{(\ell)}\boldsymbol{\eta}^{(g)} - \boldsymbol{e}^{(\ell)(g)} = 0$$


```{r}
eta_old <- matrix(rpois(m*(p+1)*K, lambda = 5), nrow = (p+1)*K, ncol = m) 
theta_old <- matrix(rpois(m*(p+1)*K, lambda = 3), nrow = (p+1)*K, ncol = m) 
alpha_old <- matrix(1, nrow = p+1, ncol = m)
Z_old <- X %*% alpha_old
e_old <- list()
for(l in 1:b) {e_old[[l]] <- Y - X %*% alpha_old - V[[l]] %*% eta_old}
u_old <- list()
for(l in 1:b) {u_old[[l]] <- matrix(1, nrow = n, ncol = m)}
w_old <- matrix(1, nrow = (p+1)*K, ncol = m)
```


# 2. Algorithm

```{r eval = FALSE}
max_iter <- 50
delta <- 0.05
lambda_1 <- 0.5
lambda_2 <- 0.5
tol_error <- 0.01

sum_V <- Reduce("+", V)
VV_prod <- lapply(V, FUN = function(x) t(x) %*% x)   # V^TV
sum_VV <- Reduce("+", VV_prod)

for(iter in 1:max_iter){
  # Process for eta
  eta_new <- matrix(nrow = (p+1)*K, ncol = m)
  Vu_prod <- mapply(function(x,y) t(x) %*% y, V, u_old, SIMPLIFY = FALSE)
  Ve_prod <- mapply(function(x,y) t(x) %*% y, V, e_old, SIMPLIFY = FALSE)
  eta_new <- (solve(sum_VV+diag(1, (p+1)*K))/delta) %*% (w_old + delta * theta_old + Reduce("+", Vu_prod)
                                                         + delta * t(sum_V) %*% (Y - Z_old)
                                                         - delta * Reduce("+", Ve_prod))
  
  # Process for theta
  theta_new <- matrix(nrow = (p+1)*K, ncol = m)
  threshold <- lambda_2/delta
  for (g in 1:m) {
    value <- eta_new[, g] - w_old[, g]/delta 
    theta_new[, g] <- case_when(value < -threshold ~ value + threshold, 
                                abs(value) < threshold ~ 0, 
                                value > threshold ~ value - threshold)
  }
  # Process for Z=XA
  Y_list <- list()
  for(i in 1:l) {Y_list[[i]] <- Y}
  VH_list <- lapply(V, FUN = function(x) x %*% eta_new)
  obj_list <- mapply(function(Y, VH, E, U) Y - VH - E - U/delta, Y_list, VH_list, e_old, u_old, SIMPLIFY = FALSE)
  obj <- Reduce("+", obj_list)/b 
  SVD <- svd(obj)
  new_singular <- sapply(SVD$d - lambda_1/(delta*b), FUN = function(x) max(x, 0))
  Z_new <- SVD$u %*% diag(new_singular) %*% t(SVD$v)
  
  # Process for e
  e_new <- list()
  for(l in 1:b){
    e_new[[l]] <- matrix(nrow = n, ncol = m)
    common_val <- Y - Z_new - V[[l]] %*% eta_new
    for(g in 1:m) {
      e_new[[l]][, g] <- case_when(e_old[[l]][, g] < 0 ~ common_val[, g] - (u_old[[l]][, g] + tau_seq[[l]] -1)/delta,
                                   e_old[[l]][, g] == 0 ~ common_val[, g] - u_old[[l]][, g]/delta, 
                                   e_old[[l]][, g] > 0 ~ common_val[, g] - (u_old[[l]][, g] + tau_seq[[l]])/delta)
    }
  }
  
  # Process for multiplier u
  u_new <- list()
  for(l in 1:b) {
    u_new[[l]] <- u_old[[l]] + delta * (Y - Z_new - V[[l]] %*% eta_new - e_new[[l]])
  }
  # Process for multiplier w
  w_new <- w_old + delta * (theta_new - eta_new)
  
  #if(e1 < tol_error) {break}
  
  eta_old <- eta_new
  theta_old <- theta_new
  alpha_old <- alpha_new
  e_old <- e_new
  u_old <- u_new
  w_old <- w_new
}
```


$\lambda_1, \lambda_2$는 임의로 0.5로 설정.


## Question
 * Stopping criteria

$$\begin{aligned}&||\eta^{k+1}-\eta^{k}||_2^2+||\theta^{k+1}-\theta^k||_2^2+||\alpha^{k+1}-\alpha^k||_2^2\\&+||e^{k+1}-e^k||_2^2+||u^{k+1}-u^k||_2^2+||w^{k+1}-w^k||_2^2 \le \epsilon_{tol}\end{aligned}$$










