---
title: "Toy Example"
author: "Sohyeon Kim"
output: 
  github_document:
    pandoc_args: --webtex
header-includes:
  - \usepackage{kotex}
---

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(splines)
library(Matrix)
library(pander)
library(gridExtra)
library(foreach)
library(doParallel)
library(Rcpp)
library(glmnet)
library(fda)

sourceCpp("functions/add_decomp_function.cpp")
source("https://raw.githubusercontent.com/S0Hye0NKim/add_decomp/master/functions/add_decomp_function.R")
```


# 0. Initial Setting

$$\begin{aligned}Y &\in \mathbb{R}^{n\times g}\\Y &= XB(\tau) + \epsilon(\tau) , \; X\sim N(0, 1)\\ X &\in\mathbb{R}^{n \times (p+1)}, \; \;B(\tau) \in \mathbb{R}^{(p+1)\times g} , \; and \; Q_\tau(\epsilon(\tau)|X) = 0\end{aligned}$$

$$\begin{aligned}\beta_0(\tau) &= \beta_0 + Q_\tau(\epsilon) \\ \epsilon(\tau) &= \epsilon- Q_\tau(\epsilon), \; \epsilon \sim N(0, 0.1^2)\end{aligned}$$

 * n : # of observation. 
 * m : # of group. 
 * p : # of covariate. 
 * r : nonzero entry in true $\Gamma(\tau)$
 * K : # of basis function.
 * b : # of regional quantiel we consider.
 
For toy example, I will use n = 300, p = 50, g = 20.

```{r}
set.seed(0)
n <- 400
m <- 10
p <- 30
#num_nz <- 15   #round((p+1)*m*(1/5)) 
b <- 31
num_rank <- 5

X <- matrix(rnorm(n*p, mean = 0, sd = 1), nrow = n) %>% cbind(1, .)   #add intercept term in X

#sp_mat <- rsparsematrix(nrow = p+1, ncol = m, nnz = num_nz)             # make sparse matrix
#for(i in 1:(p+1)) {
  #for(j in 1:m) {
    #if(abs(sp_mat[i, j]) != 0) {sp_mat[i, j] <- rnorm(1, mean = 5, sd = 0.1)}
  #}
#}
col_ind <- sample(1:m, size = m, replace = FALSE)
row_ind <- sample(1:(p+1), size = m)
sp_mat <- matrix(0, nrow = p+1, ncol = m)
for(i in 1:m) {
  sp_mat[row_ind[i], col_ind[i]] <- rnorm(1, mean = 5, sd = 0.1)
}
num_zero <- which(sp_mat==0, arr.ind = TRUE) %>% nrow
num_nz <- (p+1)*m - num_zero

LR_mat <- matrix(rnorm((p+1)*m, mean = 1, sd = 0.1), ncol = m) # make low rank matrix using SVD
SVD <- svd(LR_mat)
D_mat <- diag(SVD$d)
idx <- (num_rank+1):m
D_mat[idx, idx] <- 0
LR_mat <- SVD$u %*% D_mat %*% t(SVD$v)          

true_B <- sp_mat + LR_mat            # B(tau) = sparse matrix + low rank matrix

eps <- matrix(rnorm(n*m, mean = 0, sd = 0.1), nrow = n)

Y <- X%*%true_B + eps
```



$B(\tau)$ = sparse matrix + low rank matrix

# 1. Preliminary
 
### 1-1. Check function

```{r}
check_ft <- function(x, tau) {
  z <- ifelse(x<0, (tau-1)*x, tau*x)
  return(z)
}
```

$$\rho_\tau(x) = (\tau - \mathbb{I}(x<0))x$$

### 1-2. Basis function


```{r}
K <- 10
tau_seq <- seq(from = 0.35, to = 0.65, length.out = b)
tau_seq_real <- tau_seq[tau_seq > 0.39 & tau_seq < 0.61]

knots_seq <- seq(min(tau_seq)- 0.02, max(tau_seq) + 0.02, length.out = K)
Phi <- fda::bsplineS(tau_seq, breaks= knots_seq, norder=2, nderiv=0, returnMatrix=FALSE)
```


$$\begin{aligned}\boldsymbol{v}_i^{(\ell)} &= (\boldsymbol{x}_i \otimes \Phi(\tau_{\ell}))^T\\ &=(\Phi(\tau_\ell)^T, x_{i1}\Phi(\tau_\ell)^T, \dots, x_{ip}\Phi(\tau_\ell)^T) \\ &\in \mathbb{R}^{(p+1)K}\end{aligned}$$

### 1-3. New design matrix for regional quantile


```{r}
V <- calc_V(X, Phi)
```




 * Dimension for V : $b \times n \times (p+1)K$

$$\boldsymbol{V}^{(\ell)} = \begin{bmatrix} v_1^{(\ell)} \\v_2^{(\ell)} \\ \vdots \\ v_n^{(\ell)}\end{bmatrix} \in \mathbb{R}^{n\times (p+1)K}$$

### 1-4. Initial value for estimator


$$\boldsymbol{\eta}^{(g)}, \boldsymbol{\theta}^{(g)}, \boldsymbol{w}^{(g)}\in \mathbb{R}^{(p+1)\times K}\\ \boldsymbol{\alpha}^{(g)} \in \mathbb{R}^{p+1} \\ \boldsymbol{e}^{(\ell)(g)}, \boldsymbol{u}^{(\ell)(g)}, \in \mathbb{R}^{n}\\subject\; to \; \boldsymbol{\eta}^{(g)} - \boldsymbol{\theta}^{(g)} = 0, \; and \;\boldsymbol{Y}^{(g)} - \boldsymbol{X}\boldsymbol{\alpha}^{(g)} - \boldsymbol{V}^{(\ell)}\boldsymbol{\eta}^{(g)} - \boldsymbol{e}^{(\ell)(g)} = 0$$


# 2. Algorithm

$$\begin{aligned}\boldsymbol{Y}&=\boldsymbol{XA}+\boldsymbol{X\Gamma}(\tau)\\&=\boldsymbol{Z}+\boldsymbol{V}_{\tau_\ell}\boldsymbol{\Theta}\end{aligned}$$


$$\hat{\Gamma}_{11}(\tau_1)=\cdots=\hat{\Gamma}_{11}(\tau_\ell)=0=\Gamma_{11}(\tau)$$


## 2-1. Initial value

```{r}
lasso_coef <- matrix(nrow = p+1, ncol = m)
for(g in 1:m) {
  cv.lasso <- cv.glmnet(x = X[, -1], y = Y[, g], alpha = 1, type.measure = "mae")
  lasso_model <- glmnet(X[, -1], Y[, g], family = "gaussian", alpha = 1, lambda = cv.lasso$lambda.min)
  lasso_coef[, g] <- c(lasso_model$a0, as.vector(lasso_model$beta))
}


theta_init <- matrix(nrow = (p+1)*K, ncol = m)
for(g in 1:m) {
  for(j in 0:p) {
    theta_init[((j*K)+1):((j+1)*K), g] <- lasso_coef[j+1, g]
  }
}

Y_modified <- Y - X%*%lasso_coef

lin_model <- lm(Y_modified~., data = data.frame(X[, -1]))
alpha_init <- lin_model$coefficients


```


```{r}
max_iter <- 300
delta <- 1
lambda_1 <- 8
lambda_2 <- 40
tol_error <- 0.1
#theta_init <- matrix(0.01, nrow = (p+1)*K, ncol = m)
#alpha_init <- matrix(0.8, nrow = p+1, ncol = m)


result <- add_decomp(delta, lambda_1, lambda_2, tol_error, max_iter, X = X, Y = Y, V = V, Phi = Phi, 
                     theta_0 = theta_init, alpha_0 = alpha_init)
```


$\lambda_1, \lambda_2, \delta$ 어떻게?


```{r}
result$iter_error %>% 
  na.omit() %>%
  data.frame %>%
  mutate(iter = 1:nrow(.)) %>%
  filter(iter > 15) %>% 
  gather(key = "estimator", value = "value", -iter) %>%
  ggplot() +
  geom_line(aes(x = iter, y = value, group = estimator, color = estimator)) +
  facet_wrap(~estimator, scales = "free_y")
```


 * Stopping criteria

$$\begin{aligned}&||\eta^{k+1}-\eta^{k}||_2^2+||\theta^{k+1}-\theta^k||_2^2+||\alpha^{k+1}-\alpha^k||_2^2\\&+||e^{k+1}-e^k||_2^2+||u^{k+1}-u^k||_2^2+||w^{k+1}-w^k||_2^2 \le \epsilon_{tol}\end{aligned}$$

# 3. Evaluation

$\Gamma(\tau)$는 $\tau$에 상관 없이 동일하므로, quantile 상관 없이 $Y\approx\hat{Y}_{\tau_\ell}$ 동일.

```{r}
Y_hat_tau <- lapply(V, FUN = function(x) x %*% result$theta + X %*% result$alpha) %>%
  `names<-`(value = tau_seq)
```

### Low rank matrix

$$\frac{\sigma_1+\cdots+\sigma_5}{\sigma_1+\cdots+\sigma_{20}}\quad\text{where }\sigma_i\text{ is a singular value}$$

```{r}
sing_val <- svd(result$alpha)$d
sum(sing_val[1:5])/sum(sing_val)
rankMatrix(result$alpha)
```
$$\frac{||\hat{A}-A||_F}{||A||_F}$$

```{r}
norm(result$alpha - LR_mat, "F")/norm(LR_mat, "F")
```


### Sparse matrix

$\hat{\gamma}_j^{(g)}(\tau_\ell)=\hat{\theta}_{j}^{(g)T}\phi_s(\tau)$



```{r}
idx_tau <- (tau_seq > 0.39 & tau_seq <= 0.61)

gamma_tau_hat <- est_gamma(Phi[idx_tau, ], result$theta)
```


### sparsity pattern check

```{r}
data.frame(c("TP", "FP"), c("FN", "TN")) %>% 
  `rownames<-`(value = c("True:non_zero", "True:zero")) %>%
  `colnames<-`(value = c("Est:non_zero", "Est:zero")) %>% pander
```

```{r}
check_sp_table(true = sp_mat, est = gamma_tau_hat, table = TRUE, tol = 0.1^5, tau_seq = tau_seq_real)
```


```{r message = FALSE}
sp_heat <- data.frame(col = rep(1:m, each = (p+1)), 
           row = rep(1:(p+1), m),
           value = as.vector(sp_mat)) %>%
  ggplot(aes(x = col, y = -row, fill = value)) +
  geom_tile(color = "grey") +
  scale_fill_gradient(low = "white", high = "blue") +
  theme_bw() +
  theme(axis.line = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    axis.title.x = element_blank(), 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank()) +
  ggtitle(expression(Gamma~"("~0.5~")")) 

gamma_heat <- data.frame(col = rep(1:m, each = (p+1)), 
           row = rep(1:(p+1), m),
           value = as.vector(gamma_tau_hat[[11]])) %>%
  mutate(value = ifelse(abs(value) < 0.1^5, 0, value)) %>%
  ggplot(aes(x = col, y = -row, fill = value)) +
  geom_tile(color = "grey") +
  scale_fill_gradient2(low = "red",mid = "white", high = "blue") +
  theme_bw() +
  theme(axis.line = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    axis.title.x = element_blank(), 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank()) +
  ggtitle(expression(Gamma~"("~tau[11]~")"))


common_zero_heat <- lapply(gamma_tau_hat, FUN = function(x) data.frame(col = rep(1:m, each = (p+1)),
                                                   row = rep(1:(p+1), m),
                                                   value = as.vector(x))) %>%
  `names<-`(tau_seq_real) %>%
  bind_rows(.id = "tau") %>%
  mutate(nonzero = ifelse(abs(value) < 0.1^5, 0, 1)) %>%
  group_by(col, row) %>%
  summarise(nonzero = mean(nonzero)) %>%
  mutate(nonzero = ifelse(nonzero == 0, 0, 1) %>% as.character) %>%
  ggplot(aes(x = col, y = -row, fill = nonzero)) +
  geom_tile(color = "grey") +
  scale_fill_manual(values = c("white", "black"), name = "entry", label = c("zero", "non-zero")) + 
  theme_bw() +
  theme(axis.line = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    axis.title.x = element_blank(), 
    axis.text.x = element_blank(), 
    axis.ticks.x = element_blank(), 
    axis.title.y = element_blank(), 
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank()) +
  ggtitle(expression(Gamma~"("~tau[l]~")")) 

grid.arrange(sp_heat, gamma_heat, nrow = 1)
grid.arrange(sp_heat, common_zero_heat, nrow = 1)
```


```{r}
idx_nz_true <- which(sp_mat != 0, arr.ind = TRUE) %>% data.frame()

idx_nz_est <- lapply(gamma_tau_hat, FUN = function(x) which(abs(x) > 0.1^5, arr.ind = TRUE) %>% as_tibble) %>%
    bind_rows (.id = "tau") %>%
    group_by(row, col) %>%
    summarise(zero = n()) 



gamma_tau_hat %>%
  lapply(FUN = function(x) x[idx_nz_true$row, idx_nz_true$col] %>% diag %>%
           data.frame(row = idx_nz_true$row, col = idx_nz_true$col, value = . )) %>%
  `names<-`(value = tau_seq_real) %>%
  bind_rows(.id = "tau") %>%
  unite(col = index, row, col, sep = ",") %>%
  mutate(index = paste0("(", index, ")"), tau = as.numeric(tau)) %>%
  ggplot() +
  geom_line(aes(x = tau, y = value, group = index, color = index), size = 1.2) +
  labs(x = expression(tau[l]), title = expression(gamma~"("~tau[l]~")"~Delta~"=[0.4,0.6]"))
```
```{r}
entry_sp_mat <- data.frame(row = rep(1:(p+1), m), col = rep(1:m, each = p+1), true = sp_mat %>% as.vector ) %>%
  unite(col = index, row, col, sep = ",") %>%
  mutate(index = paste0("(", index, ")"))

gamma_tau_hat %>%
  lapply(FUN = function(x) x[idx_nz_true$row, idx_nz_true$col] %>% diag %>%
           data.frame(row = idx_nz_true$row, col = idx_nz_true$col, est = . )) %>%
  `names<-`(value = tau_seq_real) %>%
  bind_rows(.id = "tau") %>%
  unite(col = index, row, col, sep = ",") %>%
  mutate(index = paste0("(", index, ")"), tau = as.numeric(tau)) %>%
  left_join(entry_sp_mat, by = "index") %>%
  gather(key = "type", value = "value", -tau, -index) %>%
  mutate(group_var = paste0(index, "_", type)) %>%
  ggplot() +
  geom_line(aes(x = tau, y = value, group = group_var, color = index, linetype = type), size = 1.2) +
  labs(x = expression(tau[l]), title = expression(gamma~"("~tau[l]~"), b = 21, "~Delta~"=[0.4,0.6]")) +
  facet_wrap(~index)
```

### Estimation error

$$\begin{aligned}EE &= \sqrt{\int_\Delta\sum_{j=1}^p\sum_{g=1}^m\bigg(\hat{\Gamma}(\tau_\ell)-\Gamma(\tau_\ell)\bigg)^2d\tau/p},\text{   where }\tau_\ell\in[0.4,0.6]\\&=\sqrt{\sum_{g=1}^m\sum_{j=1}^p\int_\Delta\bigg(\hat{\Gamma}(\tau_\ell)-\Gamma(\tau_\ell)\bigg)^2d\tau/p}\end{aligned}$$

```{r}
h <- tau_seq_real[2] - tau_seq_real[1]

EE_mat <-  lapply(gamma_tau_hat[-1], FUN = function(x) (x - sp_mat)*(x - sp_mat) * h) %>%
  Reduce("+", .) 

EE_mat %>% apply(2, mean) %>% sum %>% sqrt

```






### Frobenius norm ratio

$$\frac{||\hat{A}-A||_F}{||A||_F}$$

```{r eval = FALSE}
lamb_1_seq <- seq(from = 0.05, to = 0.3, by = 0.05)
simul_lamb1 <- list()

for(idx in 1:length(lamb_1_seq)) {
  simul_lamb1[[idx]] <- add_decomp(delta = 0.5, lambda_1 = lamb_1_seq[idx], lambda_2 = 2.5,
                                   tol_error = 0.1, max_iter = 50)
  if((idx %% 2) == 1) print(paste0("iter = ", idx))
}
```

```{r eval = FALSE}
alpha_hat <- simul_lamb1 %>% lapply(FUN = function(x) x$alpha)

alpha_hat %>%
  lapply(FUN = function(x) norm(LR_mat - x, type = "F")/norm(LR_mat, type = "F") %>% data.frame(value = .)) %>%
  `names<-`(value = lamb_1_seq) %>%
  bind_rows(.id = "lambda_1") %>%
  ggplot() +
  geom_line(aes(x = as.numeric(lambda_1), y = value)) +
  labs(title = expression("Frobenius norm ratio when "~delta~"=0.1, "~lambda[2]~"=0.5, "~lambda[1]~"in (0.05,0.3)"), 
       x = expression(lambda[1]))
```



```{r eval = FALSE}
lapply(alpha_hat, FUN = function(x) svd(x) %>% .$d) %>%
  lapply(FUN = function(x) data.frame(ratio = (sum(x[1:num_rank])/sum(x)), rank = sum(abs(x) > 0.1^5))) %>%
  `names<-`(value = lamb_1_seq) %>%
  bind_rows(.id = "lambda1")
```



### ROC curve

```{r eval = FALSE}
lamb_2_seq <- c(seq(from = 0.1, to = 1.5, by =0.1), 1.75, 2)
simul <- list()

for(idx in 1:length(lamb_2_seq)) {
  simul[[idx]] <- add_decomp(delta = 0.1, lambda_1 = 0.2, lambda_2 = lamb_2_seq[idx], tol_error = 0.1, max_iter = 50)
  if((idx %% 2) == 1) print(paste0("iter = ", idx))
}

```


```{r}
cores <- detectCores()
cl <- makeCluster(cores[1] - 2)
registerDoParallel(cl)

simul <- foreach(lamb_2_seq = c(1, 2)) %dopar% {
  library(tidyverse)
  add_decomp_r(delta, lambda_1, lamb_2_seq, tol_error, max_iter, theta_init, alpha_init)
}

stopCluster(cl)
```




```{r eval = FALSE}
theta_hat <- simul %>%
  lapply(FUN = function(x) x$theta)

gamma_tau_hat <- list()
for(idx in 1:length(lamb_2_seq)) {
  gamma_tau_hat[[idx]] <- list()
  for(l in 1:b) {
    phi_tau <- Phi[l, ]
    gamma_tau_hat[[idx]][[l]] <- matrix(nrow = (p+1), ncol = m)
    for(i in 1:(p+1)) {
      for(j in 1:m) {
        theta_j_g <- theta_hat[[idx]][(1+(i-1)*K):(K*i), j]
        gamma_tau_hat[[idx]][[l]][i, j] <- theta_j_g %*% phi_tau
      }
    }
  }
}

ROC_table <- list()
for(idx in 1:length(lamb_2_seq)) {
  ROC_table[[idx]] <- check_sp_table(true = sp_mat, est = gamma_tau_hat[[idx]]) 
} 

ROC_table %>%
  `names<-`(value = lamb_2_seq) %>%
  bind_rows(.id = "lamb_2") %>%
  #filter(!(lamb_2 %in% c(0.55, 0.575, 0.6, 0.7, 0.8, 0.9, 1))) %>%
  arrange(FPR, TPR) %>%
  ggplot() +
  geom_line(aes(x = FPR, y = TPR)) +
  geom_point(aes(x = FPR, y = TPR)) +
  geom_abline(intercept = 0, slope = 1) +
  coord_fixed(ratio = 1) +
  ggtitle(label = expression(delta ~ " =0.1, "~lambda[1] ~ "=0.2, "~lambda[2] ~ "in (0.1, 1.5), max_iter = 50")) +
  scale_y_continuous(limits = c(0, 1))
```



```{r}
alpha_hat <- lapply(simul, FUN = function(x) x$alpha) 

lapply(alpha_hat, FUN = function(x) data.frame(FR_ratio = norm(x - LR_mat, "F")/norm(LR_mat, "F"))) %>%
  `names<-`(value = lamb_2_seq) %>%
  bind_rows(.id = "lambda_2") %>%
  mutate(lambda_2 = as.numeric(lambda_2)) %>%
  ggplot() +
  geom_line(aes(x = lambda_2, y = FR_ratio)) +
  ggtitle(label = expression(delta ~ " =0.1, "~lambda[1] ~ "=0.2, "~lambda[2] ~ "in (0.1, 2), max_iter = 50"))+
  coord_fixed(ratio = 2) 
```








