---
title: "Toy Example"
author: "Sohyeon Kim"
date: "7/21/2020"
output: 
  github_document:
    pandoc_args: --webtex
header-includes:
  - \usepackage{kotex}
---

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(splines)
library(Matrix)
library(pander)
```


# 0. Initial Setting

$$\begin{aligned}Y &\in \mathbb{R}^{n\times g}\\Y &= XB(\tau) + \epsilon(\tau) , \; X\sim N(0, 1)\\ X &\in\mathbb{R}^{n \times (p+1)}, \; \;B(\tau) \in \mathbb{R}^{(p+1)\times g} , \; and \; Q_\tau(\epsilon(\tau)|X) = 0\end{aligned}$$

$$\begin{aligned}\beta_0(\tau) &= \beta_0 + Q_\tau(\epsilon) \\ \epsilon(\tau) &= \epsilon- Q_\tau(\epsilon), \; \epsilon \sim N(0, 0.1^2)\end{aligned}$$

 * n : # of observation. 
 * m : # of group. 
 * p : # of covariate. 
 * r : nonzero entry in true $\Gamma(\tau)$
 * K : # of basis function.
 * b : # of regional quantiel we consider.
 
For toy example, I will use n = 300, p = 50, g = 20.

```{r}
set.seed(0)
n <- 300
m <- 15
p <- 20
num_nz <- round((p+1)*m*(1/3)) 
b <- 6
num_rank <- 5

X <- matrix(rnorm(n*p, mean = 0, sd = 1), nrow = n) %>% cbind(1, .)   #add intercept term in X
sp_mat <- rsparsematrix(nrow = p+1, ncol = m, nnz = num_nz)             # make sparse matrix
LR_mat <- matrix(rnorm(m*(p+1), mean = 0, sd = 1), ncol = m)      # make low rank matrix using SVD
SVD <- svd(LR_mat)
D_mat <- diag(SVD$d)
idx <- (num_rank+1):m
D_mat[idx, idx] <- 0
LR_mat <- SVD$u %*% D_mat %*% t(SVD$v)          

true_B <- sp_mat + LR_mat            # B(tau) = sparse matrix + low rank matrix

eps <- matrix(rnorm(n*m, mean = 0, sd = 0.1), nrow = n)

Y <- X%*%true_B + eps
```

$B(\tau)$ = sparse matrix + low rank matrix

# 1. Preliminary
 
### 1-1. Check function

```{r}
check_ft <- function(x, tau) {
  z <- ifelse(x<0, (tau-1)*x, tau*x)
  return(z)
}
```

$$\rho_\tau(x) = (\tau - \mathbb{I}(x<0))x$$

### 1-2. Basis function

```{r}
K <- 15

tau_seq <- seq(from = 0.4, to = 0.6, length.out = b) 
Phi <- bs(tau_seq, df = K, degree = 3, intercept = TRUE)

attr(Phi, "knots")
```


$$\begin{aligned}\boldsymbol{v}_i^{(\ell)} &= (\boldsymbol{x}_i \otimes \Phi(\tau_{\ell}))^T\\ &=(\Phi(\tau_\ell)^T, x_{i1}\Phi(\tau_\ell)^T, \dots, x_{ip}\Phi(\tau_\ell)^T) \\ &\in \mathbb{R}^{(p+1)K}\end{aligned}$$

### 1-3. New design matrix for regional quantile

```{r}
V <- list()
for(l in 1:b){
  V_l <- matrix(nrow = n, ncol = (p+1)*K)
  for(i in 1:n) {
    v_i_l <- kronecker(X[i, ], Phi[l, ])
    V_l[i, ] <- v_i_l 
  }
  V[[l]] <- V_l
}
```


 * Dimension for V : $b \times n \times (p+1)K$

$$\boldsymbol{V}^{(\ell)} = \begin{bmatrix} v_1^{(\ell)} \\v_2^{(\ell)} \\ \vdots \\ v_n^{(\ell)}\end{bmatrix} \in \mathbb{R}^{n\times (p+1)K}$$

### 1-4. Initial value for estimator


$$\boldsymbol{\eta}^{(g)}, \boldsymbol{\theta}^{(g)}, \boldsymbol{w}^{(g)}\in \mathbb{R}^{(p+1)\times K}\\ \boldsymbol{\alpha}^{(g)} \in \mathbb{R}^{p+1} \\ \boldsymbol{e}^{(\ell)(g)}, \boldsymbol{u}^{(\ell)(g)}, \in \mathbb{R}^{n}\\subject\; to \; \boldsymbol{\eta}^{(g)} - \boldsymbol{\theta}^{(g)} = 0, \; and \;\boldsymbol{Y}^{(g)} - \boldsymbol{X}\boldsymbol{\alpha}^{(g)} - \boldsymbol{V}^{(\ell)}\boldsymbol{\eta}^{(g)} - \boldsymbol{e}^{(\ell)(g)} = 0$$


```{r}
eta_old <- matrix(1, nrow = (p+1)*K, ncol = m) 
theta_old <- eta_old
alpha_old <- matrix(0, nrow = p+1, ncol = m)
Z_old <- X %*% alpha_old
e_old <- list()
for(l in 1:b) {e_old[[l]] <- Y - Z_old - V[[l]] %*% eta_old}
u_old <- list()
for(l in 1:b) {u_old[[l]] <- matrix(0, nrow = n, ncol = m)}
w_old <- matrix(0, nrow = (p+1)*K, ncol = m)
```





# 2. Algorithm

$$\begin{aligned}\boldsymbol{Y}&=\boldsymbol{XA}+\boldsymbol{X\Gamma}(\tau)\\&=\boldsymbol{Z}+\boldsymbol{V}_{\tau_\ell}\boldsymbol{\Theta}\end{aligned}$$


```{r}
add_decomp <- function(delta, lambda_1, lambda_2, tol_error, max_iter) {
  # delta = step size
  # lambda_1 = low rank penalty
  # lambda_2 = sparse penalty
  
  iter_error <- matrix(ncol = 6, nrow = max_iter) %>%
    `colnames<-`(value = c("eta", "theta", "alpha", "e", "u", "w"))
  
  sum_V <- Reduce("+", V)
  VV_prod <- lapply(V, FUN = function(x) t(x) %*% x)   # V^TV
  sum_VV <- Reduce("+", VV_prod)
  
  for(iter in 1:max_iter) {
    # Process for eta
    eta_new <- matrix(nrow = (p+1)*K, ncol = m)
    Vu_prod <- mapply(function(x,y) t(x) %*% y, V, u_old, SIMPLIFY = FALSE)
    Ve_prod <- mapply(function(x,y) t(x) %*% y, V, e_old, SIMPLIFY = FALSE)
    eta_new <- (solve(sum_VV+diag(1, (p+1)*K))/delta) %*% (w_old + delta * theta_old + Reduce("+", Vu_prod)
                                                           + delta * t(sum_V) %*% (Y - Z_old)
                                                           - delta * Reduce("+", Ve_prod))
    
    # Process for theta
    theta_new <- matrix(nrow = (p+1)*K, ncol = m)
    for (g in 1:m) {
      r_g <- eta_new[, g] - w_old[, g]/delta
      value <- 1 - (lambda_2/(delta * abs(r_g)))
      theta_new[, g] <- ifelse(value > 0, value * r_g, 0)
    }
    
    # Process for Z=XA
    Y_list <- list()
    for(i in 1:l) {Y_list[[i]] <- Y}
    VH_list <- lapply(V, FUN = function(x) x %*% eta_new)
    obj_list <- mapply(function(Y, VH, E, U) Y - VH - E - U/delta, Y_list, VH_list, e_old, u_old, SIMPLIFY = FALSE)
    obj <- Reduce("+", obj_list)/b 
    SVD <- svd(obj)
    new_singular <- sapply(SVD$d - lambda_1/(delta*b), FUN = function(x) max(x, 0))
    Z_new <- SVD$u %*% diag(new_singular) %*% t(SVD$v)
    alpha_new <- solve(t(X) %*% X) %*% t(X) %*% Z_new
    
    # Process for e
    e_new <- list()
    for(l in 1:b){
      e_new[[l]] <- matrix(nrow = n, ncol = m)
      for(g in 1:m) {
        error <- Y[, g] - Z_new[, g] - V[[l]] %*% eta_new[, g]   #error = Y - XA - VH
        value <- error + u_old[[l]][, g]/delta
        e_new[[l]][, g] <- case_when(value > tau_seq[l]/(n*delta) ~ value - tau_seq[l]/(n*delta), 
                                     value < (tau_seq[l]-1)/(n*delta) ~ value - (tau_seq[l]-1)/(n*delta), 
                                     value >=(tau_seq[l]-1)/(n*delta) & value <= tau_seq[l]/(n*delta) ~ 0)
      }
    }
    
    # Process for multiplier u
    u_new <- list()
    for(l in 1:b) {
      u_new[[l]] <- u_old[[l]] + delta * (Y - Z_new - V[[l]] %*% eta_new - e_new[[l]])
    }
    
    # Process for multiplier w
    w_new <- w_old + delta * (theta_new - eta_new)
    
    # Update iteration error
    iter_error[iter, "eta"] <- Matrix::norm(eta_old - eta_new, type = "F")
    iter_error[iter, "theta"] <- Matrix::norm(theta_old - theta_new, type = "F")
    iter_error[iter, "alpha"] <- Matrix::norm(alpha_old - alpha_new, type = "F")
    e_diff <- mapply(FUN = function(old, new) old - new, e_old, e_new, SIMPLIFY = FALSE)  # sum of frobenius norm
    iter_error[iter, "e"] <- lapply(e_diff, FUN = function(x) Matrix::norm(x, type = "F")) %>% Reduce("+", .)
    u_diff <- mapply(FUN = function(old, new) old - new, u_old, u_new, SIMPLIFY = FALSE)
    iter_error[iter, "u"] <- lapply(u_diff, FUN = function(x) Matrix::norm(x, type = "F")) %>% Reduce("+", .)
    iter_error[iter, "w"] <- Matrix::norm(w_old - w_new, type = "F")
    
    if(sum(iter_error[iter, ]) < tol_error) break
    
    eta_old <- eta_new
    theta_old <- theta_new
    Z_old <- Z_new
    alpha_old <- alpha_new
    e_old <- e_new
    u_old <- u_new
    w_old <- w_new
  }
  
  return(list(eta = eta_new, 
              theta = theta_new, 
              alpha = alpha_new, 
              e = e_new, 
              u = u_new, 
              w = w_new, 
              iter_error = iter_error))
}
```


```{r}
max_iter <- 50
delta <- 0.25
lambda_1 <- 0.2
lambda_2 <- 1
tol_error <- 0.1

result <- add_decomp(delta, lambda_1, lambda_2, tol_error, max_iter)
```



$\lambda_1, \lambda_2, \delta$ 어떻게?


```{r}
result$iter_error %>% 
  na.omit() %>%
  data.frame %>%
  mutate(iter = 1:nrow(.)) %>%
  #filter(iter > 15) %>% 
  gather(key = "estimator", value = "value", -iter) %>%
  ggplot() +
  geom_line(aes(x = iter, y = value, group = estimator, color = estimator)) +
  facet_wrap(~estimator, scales = "free_y")
```


 * Stopping criteria

$$\begin{aligned}&||\eta^{k+1}-\eta^{k}||_2^2+||\theta^{k+1}-\theta^k||_2^2+||\alpha^{k+1}-\alpha^k||_2^2\\&+||e^{k+1}-e^k||_2^2+||u^{k+1}-u^k||_2^2+||w^{k+1}-w^k||_2^2 \le \epsilon_{tol}\end{aligned}$$

# 3. Evaluation

$\Gamma(\tau)$는 $\tau$에 상관 없이 동일하므로, quantile 상관 없이 $Y\approx\hat{Y}_{\tau_\ell}$ 동일.

```{r}
Y_hat_tau <- lapply(V, FUN = function(x) x %*% result$theta + X %*% result$alpha) %>%
  `names<-`(value = tau_seq)
```

### Low rank matrix

$$\frac{\sigma_1+\cdots+\sigma_5}{\sigma_1+\cdots+\sigma_{20}}\quad\text{where }\sigma_i\text{ is a singular value}$$

```{r}
sing_val <- svd(result$alpha)$d
sum(sing_val[1:5])/sum(sing_val)
rankMatrix(result$alpha)
```





### Sparse matrix

$\hat{\gamma}_j^{(g)}(\tau_\ell)=\hat{\theta}_{j}^{(g)T}\phi_s(\tau)$


```{r}
gamma_tau_hat <- list()
for(l in 1:b) {
  phi_tau <- Phi[l, ]
  gamma_tau_hat[[l]] <- matrix(nrow = (p+1), ncol = m)
  for(i in 1:(p+1)) {
    for(j in 1:m) {
      theta_j_g <- result$theta[(1+(i-1)*15):(15*i), j]
      gamma_tau_hat[[l]][i, j] <- theta_j_g %*% phi_tau
    }
  }
}
```


### sparsity pattern check

```{r}
data.frame(c("TP", "FP"), c("FN", "TN")) %>% 
  `rownames<-`(value = c("True:non_zero", "True:zero")) %>%
  `colnames<-`(value = c("Est:non_zero", "Est:zero")) %>% pander
```


```{r}
num_zero <- which(sp_mat==0, arr.ind = TRUE) %>% nrow
print(paste0("The number of zero entry in sp_mat is ", num_zero))
```

```{r}
check_sp_table <- function(true, est, tol = 0.1^5, table = FALSE, nnz = num_nz, nz = num_zero) {
  zero_idx_true <- which(abs(true) < tol, arr.ind = TRUE) %>% as_tibble
  zero_idx_est <- which(abs(est) < tol, arr.ind = TRUE) %>% as_tibble
  TN <- semi_join(zero_idx_true, zero_idx_est, by = c("row", "col"))
  FP <- anti_join(zero_idx_true, zero_idx_est, by = c("row", "col"))
  FN <- anti_join(zero_idx_est, zero_idx_true, by = c("row", "col"))
  result <- data.frame(Positive = c(nnz - nrow(FN), nrow(FP)), Negative = c(nrow(FN), nrow(TN))) %>%
                             `rownames<-`(value = c("Positive", "Negative"))
  if(table == FALSE) {return(data.frame(FPR = result[2, 1]/nz, TPR = result[1, 1]/nnz))
    } else {return(result)}
}
```


```{r}
lapply(gamma_tau_hat, FUN = function(x) check_sp_table(true = sp_mat, est = x, table = FALSE)) %>%
  bind_rows %>%
  summarise_all(mean)
```



### ROC curve

```{r eval = FALSE}
lamb_2_seq <- seq(from = 0.1, to = 1.1, by = 0.1)
result <- list()

for(idx in 1:length(lamb_2_seq)) {
  result[[idx]] <- add_decomp(delta = 0.1, lambda_1 = 0.2, lambda_2 = lamb_2_seq[idx], tol_error = 0.1, max_iter = 50)
  if((idx %% 2) == 1) print(paste0("iter = ", idx))
}
```


```{r eval = FALSE}
theta_hat <- result %>%
  lapply(FUN = function(x) x$theta)

gamma_tau_hat <- list()
for(idx in 1:length(lamb_2_seq)) {
  gamma_tau_hat[[idx]] <- list()
  for(l in 1:b) {
    gamma_tau_hat[[idx]][[l]] <- matrix(nrow = (p+1), ncol = m)
    for(i in 1:(p+1)) {
      for(j in 1:m) {
        theta_j_g <- theta_hat[[idx]][(1+(i-1)*K):(K*i), j]
        gamma_tau_hat[[idx]][[l]][i, j] <- theta_j_g %*% phi_tau
      }
    }
  }
}

ROC_table <- list()
for(idx in 1:length(lamb_2_seq)) {
  ROC_table[[idx]] <-gamma_tau_hat[[idx]] %>%
    lapply(FUN = function(x) check_sp_table(true = sp_mat, est = x)) %>%
    bind_rows(.id = "tau") %>%
    mutate(tau = tau_seq)
}

ROC_table %>%
  `names<-`(value = lamb_2_seq) %>%
  bind_rows(.id = "lamb_2") %>%
  mutate(tau = as.factor(tau)) %>% 
  ggplot() +
  geom_line(aes(x = FPR, y = TPR, group = tau, color = tau)) +
  geom_point(aes(x = FPR, y = TPR))
```















